\documentclass[12pt]{article}
\usepackage[dvipsnames]{xcolor}
\usepackage{mathtools,amssymb}
\usepackage[margin=.65in]{geometry}
\pagestyle{empty}
\everymath{\displaystyle}
\usepackage{tabularx,ragged2e,booktabs,caption}
\usepackage{upquote}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta}
\usepackage{array}
\usepackage{parskip}
\usepackage[colorlinks=true, urlcolor=blue]{hyperref}
\usepackage{fancyvrb}
\fvset{xleftmargin=10mm}

\renewcommand{\familydefault}{\sfdefault}
\usepackage{mathptmx} %Times math and text font
\usepackage{helvet}
\usepackage{mathastext}

% groovy underlining
\usepackage[normalem]{ulem}
\usepackage{contour}
\renewcommand{\ULdepth}{1.8pt}
\contourlength{0.8pt}
\newcommand{\myuline}[1]{%
  \uline{\phantom{#1}}%
  \llap{\contour{white}{#1}}%
}

% custom commands, envirnments, lists and column types
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{titleps} % alternative to fncyhdr
\newpagestyle{main}{%
  \setheadrule{0pt} %
  \sethead[][][]{}{}{}% \sethead{<left>}{<centre>}{<right>}
  \setfoot[][][]{}{}{}% \setfoot{<left>}{<centre>}{<right>}
}
\pagestyle{main}







\begin{document}
\raggedright

\renewcommand{\ULdepth}{2.8pt}
\renewcommand{\ULthickness}{1.8pt}

\myuline{\makebox[\textwidth][l]{\LARGE Git and GitHub Notes}}

\vspace{-1ex}
\subsection*{RESOURCES}
\begin{itemize}[itemsep=-.5ex]
\item Videos:
\begin{itemize}[itemsep=-.5ex]
\item Kevin Stravert: \href{https://www.youtube.com/watch?v=tRZGeaHPoaw}{https://www.youtube.com/watch?v=tRZGeaHPoaw}
\item Gwen Faraday:
\href{https://www.youtube.com/watch?v=-9Tb76tBoFU}{https://www.youtube.com/watch?v=-9Tb76tBoFU}
\end{itemize}
\item Official Git web site: \href{https://www.git-scm.com/}{https://www.git-scm.com/}
\item Official GitHub.com web site: \href{https://github.com/}{https://github.com/}
\item Git and GitHub.com cheat sheet: \href{https://education.github.com/git-cheat-sheet-education.pdf}{https://education.github.com/git-cheat-sheet-education.pdf}
\item Git Reference Manual: \href{https://git-scm.com/docs}{https://git-scm.com/docs}
\item Git Overview Book: \href{https://git-scm.com/book/en/v2}{https://git-scm.com/book/en/v2}
\item Sample ignore files: \href{https://github.com/github/gitignore}{https://github.com/github/gitignore}
\item Hyper.is Terminal: \href{https://hyper.is/}{https://hyper.is/}
\item A decent Git command reference: \href{https://dev.classmethod.jp/articles/git-bash-commands/}{https://dev.classmethod.jp/articles/git-bash-commands/}
\end{itemize}

\section{Install Git and Git Bash}

\subsection*{Windows}
Download from \href{https://git-scm.com/}{https://git-scm.com/}\par
Select all of the default settings. Git Bash has also been installed and can be used instead of the command prompt. Git Bash supports UNIX commands for Git.

\subsection*{Linux}
\begin{Verbatim}
sudo apt-get update
apt-get install git
\end{Verbatim}
\subsection*{After installation}
Check the installed version.
\begin{Verbatim}
git --version
\end{Verbatim}
Set up your username and email address.
\begin{Verbatim}
git config --global user.email jamorgan75@protonmail.com
git config --global user.name "James Morgan"
\end{Verbatim}
Set default branch name to ``main":
\begin{Verbatim}
git config --global init.default branch main
\end{Verbatim}
View configuration.
\begin{Verbatim}
git config --list
\end{Verbatim}


\newpage


\section{Get help}
Ask git for help:
\begin{Verbatim}
git config -h
\end{Verbatim}
If you want more detailed information,
\begin{Verbatim}
git help config
\end{Verbatim}
This follows the format \verb|git help <command>|





\section{Start a New Project}
\subsection*{Using Github}
Create a new repository on \href{https://github.com/}{Github}. If you are planning on publishing this repository as the main page of your Github pages, the repository should be named\; \verb|<username>.github.io|\par
From within Github.com, click on \verb|Code| and copy the Github URL. Navigate to the directory that the repository will be cloned to locally. Then type the following and paste the address.
\begin{Verbatim}
git clone <URL>
\end{Verbatim}


\subsection*{Local Only}
Within terminal, change directory to the folder for which you wish to create a repository. Type the following and the press Enter:
\begin{Verbatim}
git init
\end{Verbatim}
This creates a hidden folder \verb|.git| within the directory that contains the repository files.

\subsection*{Work On Your Code}
Open your editor/IDE, create necessary files and start coding.






\newpage





\section{Working Towards the Initial Commit}
\subsection*{Git status}
We can find the status of our repository with the following command:
\begin{Verbatim}
git status
\end{Verbatim}
This command will return the following information:
\begin{itemize}[nosep]
\item branch name
\item commits
\item tracked (green) or untracked (red) files
\end{itemize}

\subsection*{Track and Untrack Files}
To track a file:
\begin{Verbatim}
git add index.htm
\end{Verbatim}
To untrack (unstage) a file:
\begin{Verbatim}
git rm --cached index.htm
git add notes_on_git_and_github.tex
\end{Verbatim}
To untrack (unstage) a file:
\begin{Verbatim}
git rm --cached notes_on_git_and_github.tex
\end{Verbatim}


\subsection*{Ignore files with {\ttfamily .gitignore}}
Create a \verb|.txt| file in the directory and add the filetype. Comment with the hashtag symbol.
\begin{Verbatim}[frame=single]
# ignore all .txt files
*.txt
\end{Verbatim}
Ignoring files is useful if you have files with sensitive information within the directory.\par
To see a comprehensive list of all the ways to ignore files, visit \href{https://github.com/github/gitignore}{https://github.com/github/gitignore}

\subsection*{Track all files / add to staging}
To track all files within the directory, use one of the following:
\begin{Verbatim}
git add --all
git add -A
git add .
\end{Verbatim}


\subsection*{Commit}
To commit is to take a snap shot of what your repository looks like at this point in time.
\begin{Verbatim}
git commit -m "first commit - committing all files to the repository"
\end{Verbatim}

\section{Change files and view differences}
If we change and save a file, the file is now categorized as ``modified.''\par
We can compare the differences in the modified file with the following command:
\begin{Verbatim}
git diff
\end{Verbatim}
The original text will be in red, and the changed text will be in green.\par
We can now use the add command, and the file will be sitting in ``staging''.\par
Within Git, we have three different environments: working files, staging, and commit.
\begin{figure}[h!]
\centering
\begin{tikzpicture}
\node[draw,align=center,minimum height = 12mm, minimum width=30mm] (working) {WORKING\\FILES};
\node[draw,align=center,minimum height = 12mm, minimum width=30mm,below = 5mm of working] (staging) {STAGING};
\node[draw,align=center,minimum height = 12mm, minimum width=30mm,below = 5mm of staging] (commit) {COMMIT};
\node[right = 5mm of working] {Files are edited here};
\node[right = 5mm of staging] {Files are held to be committed};
\node[right = 5mm of commit] {A record of change is created for a file};
\end{tikzpicture}
\end{figure}
\par
To remove files from staging:
\begin{Verbatim}
git restore --staged index.htm
\end{Verbatim}
If we commit now, the \verb|index.htm| file will not be included as it is in the working environment.

\begin{figure}[h!]
\centering
\begin{tikzpicture}[rounded corners,-{Stealth[length=6pt]},nodes={font=\footnotesize}]
\node[fill=Bittersweet!25,draw, align=center,minimum height = 12mm, minimum width=30mm] (working) {Working\\directory};
\node[fill=Bittersweet!50,draw,align=center,minimum height = 12mm, minimum width=30mm,above right = 5mm and 5mm of working] (staging) {Staging\\index};
\node[fill=Bittersweet!75,draw,align=center,minimum height = 12mm, minimum width=30mm,above right = 5mm and 5mm of staging] (local) {Local\\repository};
\node[fill=Bittersweet,draw,align=center,minimum height = 12mm, minimum width=30mm,above right = 5mm and 5mm of local] (remote) {Remote\\repository};

\draw (working.130) --++ (90:22mm) -| (staging.150) node[pos=.05,anchor=south west]{add};
\draw (staging.130) --++ (90:22mm) -| (local.150) node[pos=.05,anchor=south west]{commit};
\draw (local.130) --++ (90:22mm) -| (remote.150) node[pos=.05,anchor=south west]{push};

\draw (remote.230) --++ (270:22mm) -| (local.310) node[pos=.05,anchor=south east]{pull};
\draw (local.230) --++ (270:39mm) -| (working.310) node[pos=.05,anchor=south east]{reset};
\draw (remote.230) --++ (270:59mm) -| (working.230) node[pos=.05,anchor=south east]{};
\draw[-] (staging.230) --++ (270:22mm) -| (working.310);
\end{tikzpicture}
\end{figure}


\section{Bypass staging and commit}
The following skips the staging steps:
\begin{Verbatim}
git commit -a -m "updated text to free range"
\end{Verbatim}
\section{Delete / remove files}
\begin{Verbatim}
git rm "secret recipe.htm"
\end{Verbatim}
If Git is tracking the deleted file, status will show that the file has been deleted.

\section{Restore files}
\begin{Verbatim}
git restore "secret recipe.htm"
\end{Verbatim}
This recovers the file.

\section{Rename files}
\begin{Verbatim}
git mv "KCC Logo.png" "Primary Logo.png"
git commit -m "chagnnnged the file name of an image"
\end{Verbatim}

\section{View commit history with git log}
\begin{Verbatim}
git log
\end{Verbatim}
To get an abbreviated view:
\begin{Verbatim}
git log --oneline
\end{Verbatim}

\section{Amend commit}
\begin{Verbatim}
git commit -m "changed file name to Primary Logo.png" --amend
\end{Verbatim}

\section{View changes in commits}
\begin{Verbatim}
git log -p
\end{Verbatim}
To find the different capabilities of \verb|git log| is to use the help feature:
\begin{Verbatim}
git help log
\end{Verbatim}


\section{Reset to previous commit}
To jump back to a previous commit:
\begin{Verbatim}
git reset c193894
\end{Verbatim}

\section{Rebase git repository}
To modify your commits:
\begin{Verbatim}
git rebase -i --root
\end{Verbatim}
To exit this menu, press :X and then Enter.


\section{Branches}
To create a new branch, from the directory we enter:
\begin{Verbatim}
git branch FixTemp
\end{Verbatim}
To see branches, type:
\begin{Verbatim}
git branch
\end{Verbatim}
We are currently in the branch with the asterisk.\par
To switch into another branch, we type:
\begin{Verbatim}
git switch FixTemp
\end{Verbatim}
Now any changes made to the files in the directory will be associated with the current branch. Our commits will be applied to that branch.\par
Switching back to main,
\begin{Verbatim}
git switch main
\end{Verbatim}
our files in the directory no reflect the changes made in the FixTemp branch. To bring the changes to the main branch, we need to merge the changes in.

\section{Merge branches}
We will need to specify which branch we want to merge with main.
\begin{Verbatim}
git merge -m "Merge fixtemp back to main" FixTemp
\end{Verbatim}

\section{Delete branch}
Now that we've merged the FixTemp branch with main, we no longer need the FixTemp branch.
\begin{Verbatim}
git branch -d FixTemp
\end{Verbatim}

\section{Merge conflicts}
If changes in main are committed before a branch is merged, we will have a merge conflict.\par
The following line will create a new branch and switch to that branch:
\begin{Verbatim}
git switch -c UpdateText
\end{Verbatim}

\section{Typical Git flow}
You have some feature or bug to work on, so you create a new branch.
\begin{enumerate}[nosep]
\item Create new branch
\item Make all changes
\item Merge into main
\item Delete branch that you were working on
\end{enumerate}


\section{Set up GitHub account}
Your GitHub account can be created at \href{https://github.com/}{https://github.com/}

\section{Create new cloud repository}
On the left-hand side of the GitHub web page, you should see the option to create a new repository (a.k.a. repo). Alternatively,  you can visit \href{github.new}{github.new}, and this will drop you on the new repo page.\par

Enter a name for the repo. If you make the repo private, you can assign different individuals access to the repo.

\section{Push local repo to GitHub}
\section{Working with files}
\section{Edit repo details}
\section{Issues}
\section{Pull requests}
\section{Actions, Projects, Wiki, Security, Insights, Settings}
\section{Releases}
\section{Fetch and pull}
\section{Wrap up}





\end{document}




